---
title: "GO MWU plotting"
author: "Jill Ashey"
date: "2024-07-23"
output: html_document
---

This script will plot the outputs from GO MWU. Specifically, I'm interested in plotting the delta rank of shared GO terms of each treatment comparison against one another. This will tell us if the GO terms that are shared across treatment comparisons have similar expression patterns. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggrepel)
library(cowplot)
library(kableExtra)
library(BiocManager)
library(patchwork)
library(csranks)
#library("GOSim")
#library(rrvgo)
#library(org.Ce.eg.db)
```

## High v Control and High v Mid 

#### Biological Processes

Read in BP data. 
```{r}
bp_hvc=read.table("GO_MWU/MWU_BP_DE_High_v_Control_LFC.csv",header=T)
bp_hvm=read.table("GO_MWU/MWU_BP_DE_High_v_Mid_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
bp_good <- intersect(bp_hvc$term, bp_hvm$term)

# Filter sets
bp_hvc_filt <- bp_hvc[bp_hvc$term %in% bp_good,]
bp_hvm_filt <- bp_hvm[bp_hvm$term %in% bp_good,]
```

There are 5751 shared GO terms between the High v Control and High v Mid comparisons

Combine sets 
```{r}
ress <- merge(bp_hvc_filt, bp_hvm_filt, by = "term")
```

The x is High v Control and the y is High v Mid. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

106 shared BP GO terms are < 0.01.

Fit a linear model to find the slope 
```{r}
# Fit a linear model
model <- lm(delta.rank.y ~ delta.rank.x, data = ress_filt)

# Extract the slope
slope <- coef(model)[2]

print(paste("The slope is:", round(slope, 4)))
```

Fit a rank regression to find the slope 
```{r}
## using csranks package
lmr_model <- lmranks(r(delta.rank.y) ~ r(delta.rank.x), data=ress_filt)
summary(lmr_model)

slope_rank <- coef(lmr_model)[2]
```

Create string for enriched delta ranks - these are the delta ranks that have the highest positive value in both comparisons 
```{r}
enrich_list <- c("single strand break repair", "G-quadruplex DNA unwinding", "V(D)J recombination", "negative regulation of telomere capping", "mRNA 3’-splice recognition")
enrich_list_label <- paste(enrich_list, collapse = "\n")
# Create a data frame for the label position
enrich_label_data <- data.frame(
  x = 800, # Adjust x position as needed
  y = 3500, # Adjust y position as needed
  label = enrich_list_label
)
```

Create string for underrepresented delta ranks - these are the delta ranks that have the most negative value in both comparisons 
```{r}
under_list <- c("lipoxygenase pathway", "axonemal dynein complex assembly", "long-chain fatty acid biosynthetic process", "cobalamin metabolic process", "fatty acid beta-oxidation")
under_list_label <- paste(under_list, collapse = "\n")
# Create a data frame for the label position
under_label_data <- data.frame(
  x = -3700, # Adjust x position as needed
  y = -650, # Adjust y position as needed
  label = under_list_label
)
```

Modify df to signify which points to color 
```{r}
ress_filt <- ress_filt %>%
  mutate(include = case_when(
    str_detect(name.x, "single strand break repair|G-quadruplex DNA unwinding|V\\(D\\)J recombination|negative regulation of telomere capping|mRNA.*3'-splice.*recognition") |
    str_detect(name.y, "single strand break repair|G-quadruplex DNA unwinding|V\\(D\\)J recombination|negative regulation of telomere capping|mRNA.*3'-splice.*recognition") ~ "enriched",
    str_detect(name.x, "lipoxygenase pathway|axonemal dynein complex assembly|long-chain fatty acid biosynthetic process|fatty acid beta-oxidation|cobalamin metabolic process") |
    str_detect(name.y, "lipoxygenase pathway|axonemal dynein complex assembly|long-chain fatty acid biosynthetic process|fatty acid beta-oxidation|cobalamin metabolic process") ~ "under",
    
    TRUE ~ NA_character_
  ))
```

Plot 
```{r}
# Plot
plot1 <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y, color = include)) +
  # Negative horizontal dashed line (blue)
  geom_segment(aes(x = -Inf, xend = 0, y = 0, yend = 0), linetype = "dashed", alpha = 0.95, size = 1, color = "blue") +
  
  # Positive horizontal dashed line (red)
  geom_segment(aes(x = 0, xend = Inf, y = 0, yend = 0), linetype = "dashed", alpha = 0.95, size = 1, color = "red") +
  
  # Negative vertical dashed line (blue)
  geom_segment(aes(x = 0, xend = 0, y = -Inf, yend = 0), linetype = "dashed", alpha = 0.95, size = 1, color = "blue") +
  
  # Positive vertical dashed line (red)
  geom_segment(aes(x = 0, xend = 0, y = 0, yend = Inf), linetype = "dashed", alpha = 0.95, size = 1, color = "red") +
  
  geom_point(size = 4, show.legend = FALSE) +
  scale_size("size") +
  scale_x_continuous(limits = c(-3700, 3700), breaks = seq(-3000, 3000, 1000)) +
  scale_y_continuous(limits = c(-3700, 3700), breaks = seq(-3000, 3000, 1000)) +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (High v. Mid)") +
  #geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.75) +
  #geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.75) +
  scale_color_manual(values = c("red", "blue", "black")) +
  # geom_label_repel(
  #   data = enrich_label_data,
  #   aes(x = x, y = y, label = label),
  #   box.padding = unit(1, "lines"),
  #   point.padding = unit(0.5, "lines"),
  #   size = 2.5,
  #   color = "red", # Text color
  #   fill = "white", # Background color of the label
  #   segment.color = NA) + # No line segment needed here
  # geom_label_repel(
  #   data = under_label_data,
  #   aes(x = x, y = y, label = label),
  #   box.padding = unit(1, "lines"),
  #   point.padding = unit(0.5, "lines"),
  #   size = 2.5,
  #   color = "blue", # Text color
  #   fill = "white", # Background color of the label
  #   segment.color = NA) + # No line segment needed here
  theme_cowplot()+
  annotate("label", label = "Overrepresented\nin High", x = 2600, y = 500, size = 7, color = "red", label.size = 0.8)+
  annotate("label", label = "Underrepresented\nin High", x = -2600, y = -500, size = 7, color = "blue", label.size = 0.8)+
  annotate("text", label = paste("Slope:", round(slope_rank, 2)), x = 2300, y = -3600, size = 6, color = "black") +
  #annotate("text", x = 700, y = 3100, label = "Overrepresented in High", size = 3, color = "red", face = "bold") +
  theme(axis.text = element_text(size = 25),
        axis.title = element_text(size = 28)); plot1

ggsave("../output/GO_MWU/HighvControl_HighvMid_BP_DeltaRank.pdf", plot = plot1, width = 30, height = 25, units = "cm")
ggsave("../output/GO_MWU/HighvControl_HighvMid_BP_DeltaRank.png", plot = plot1, width = 30, height = 25, units = "cm")
```



#### Molecular Function 

Read in MF data. 
```{r}
mf_hvc=read.table("GO_MWU/MWU_MF_DE_High_v_Control_LFC.csv",header=T)
mf_hvm=read.table("GO_MWU/MWU_MF_DE_High_v_Mid_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
mf_good <- intersect(mf_hvc$term, mf_hvm$term)

# Filter sets
mf_hvc_filt <- mf_hvc[mf_hvc$term %in% mf_good,]
mf_hvm_filt <- mf_hvm[mf_hvm$term %in% mf_good,]
```

There are 1105 shared GO terms between the High v Control and High v Mid comparisons

Combine sets 
```{r}
ress <- merge(mf_hvc_filt, mf_hvm_filt, by = "term")
```

The x is High v Control and the y is High v Mid. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

28 shared MF GO terms are < 0.01.

Fit a linear model to find the slope 
```{r}
# Fit a linear model
model <- lm(delta.rank.y ~ delta.rank.x, data = ress_filt)

# Extract the slope
slope <- coef(model)[2]

print(paste("The slope is:", round(slope, 4)))
```

Fit a rank regression to find the slope 
```{r}
## using csranks package
lmr_model <- lmranks(r(delta.rank.y) ~ r(delta.rank.x), data=ress_filt)
summary(lmr_model)

slope_rank <- coef(lmr_model)[2]
```

Create string for enriched delta ranks - these are the delta ranks that have the highest positive value in both comparisons 
```{r}
enrich_list <- c("single-stranded DNA helicase activity", "DNA phosphodiesterase activity", "histone kinase activity", "single-stranded DNA binding", "DNA helicase activity")
enrich_list_label <- paste(enrich_list, collapse = "\n")
# Create a data frame for the label position
enrich_label_data <- data.frame(
  x = 1000, # Adjust x position as needed
  y = 3300, # Adjust y position as needed
  label = enrich_list_label
)
```

Create string for underrepresented delta ranks - these are the delta ranks that have the most negative value in both comparisons 
```{r}
under_list <- c("oxidoreductase activity", "oxidoreductase activity, acting on single donors with incorporation of molecular oxygen")
under_list_label <- paste(under_list, collapse = "\n")
# Create a data frame for the label position
under_label_data <- data.frame(
  x = -3500, # Adjust x position as needed
  y = -1000, # Adjust y position as needed
  label = under_list_label
)
```

Color points to be labeled on plot 
```{r}
ress_filt <- ress_filt %>%
  mutate(include = case_when(
    str_detect(name.x, "single-stranded DNA helicase activity|DNA phosphodiesterase activity|histone kinase activity|single-stranded DNA binding|DNA helicase activity") |
    str_detect(name.y, "single-stranded DNA helicase activity|DNA phosphodiesterase activity|histone kinase activity|single-stranded DNA binding|DNA helicase activity") ~ "enriched",
    str_detect(name.x, "oxidoreductase activity|oxidoreductase activity, acting on single donors with incorporation of molecular oxygen") |
    str_detect(name.y, "oxidoreductase activity|oxidoreductase activity, acting on single donors with incorporation of molecular oxygen") ~ "under",
    
    TRUE ~ NA_character_
  ))
```

Plot 
```{r}
# Plot
plot2 <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y, color = include)) +
    # Negative horizontal dashed line (blue)
  geom_segment(aes(x = -Inf, xend = 0, y = 0, yend = 0), linetype = "dashed", alpha = 0.95, size = 1, color = "blue") +
  
  # Positive horizontal dashed line (red)
  geom_segment(aes(x = 0, xend = Inf, y = 0, yend = 0), linetype = "dashed", alpha = 0.95, size = 1, color = "red") +
  
  # Negative vertical dashed line (blue)
  geom_segment(aes(x = 0, xend = 0, y = -Inf, yend = 0), linetype = "dashed", alpha = 0.95, size = 1, color = "blue") +
  
  # Positive vertical dashed line (red)
  geom_segment(aes(x = 0, xend = 0, y = 0, yend = Inf), linetype = "dashed", alpha = 0.95, size = 1, color = "red") +
  
  geom_point(size = 4, show.legend = FALSE) +
  scale_size("size") +
  scale_x_continuous(limits = c(-3700, 3700), breaks = seq(-3000, 3000, 1000)) +
  scale_y_continuous(limits = c(-3700, 3700), breaks = seq(-3000, 3000, 1000)) +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (High v. Mid)") +
  #geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.75) +
  #geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.75) +
  scale_color_manual(values = c("red", "blue", "black")) +
  # geom_label_repel(
  #   data = enrich_label_data,
  #   aes(x = x, y = y, label = label),
  #   box.padding = unit(1, "lines"),
  #   point.padding = unit(0.5, "lines"),
  #   size = 2.5,
  #   color = "red", # Text color
  #   fill = "white", # Background color of the label
  #   segment.color = NA) + # No line segment needed here
  # geom_label_repel(
  #   data = under_label_data,
  #   aes(x = x, y = y, label = label),
  #   box.padding = unit(1, "lines"),
  #   point.padding = unit(0.5, "lines"),
  #   size = 2.5,
  #   color = "blue", # Text color
  #   fill = "white", # Background color of the label
  #   segment.color = NA) + # No line segment needed here
  theme_cowplot() +
  annotate("label", label = "Overrepresented\nin High", x = 2600, y = 500, size = 7, color = "red", label.size = 0.8)+
  annotate("label", label = "Underrepresented\nin High", x = -2600, y = -500, size = 7, color = "blue", label.size = 0.8)+
  annotate("text", label = paste("Slope:", round(slope_rank, 2)), x = 2300, y = -3600, size = 6, color = "black") +
  theme(axis.text = element_text(size = 25),
        axis.title = element_text(size = 28)); plot2

ggsave("../output/GO_MWU/HighvControl_HighvMid_MF_DeltaRank.pdf", plot = plot2, width = 30, height = 25, units = "cm")
ggsave("../output/GO_MWU/HighvControl_HighvMid_MF_DeltaRank.png", plot = plot2, width = 30, height = 25, units = "cm")
```

Make plots into one figure and save 
```{r}
plot <- plot_grid(plot1, plot2, labels = c('A', 'B'))

ggsave("../output/GO_MWU/HighvMid_HighvControl_DeltaRank.pdf", plot = plot, width = 45, height = 15, units = "cm")
ggsave("../output/GO_MWU/HighvMid_HighvControl_DeltaRank.png", plot = plot, width = 45, height = 15, units = "cm")
```


## High v Control and Mid v Control

#### Biological Processes 

Read in BP data. 
```{r}
bp_hvc=read.table("GO_MWU/MWU_BP_DE_High_v_Control_LFC.csv",header=T)
bp_mvc=read.table("GO_MWU/MWU_BP_DE_Mid_v_Control_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
bp_good <- intersect(bp_hvc$term, bp_mvc$term)

# Filter sets
bp_hvc_filt <- bp_hvc[bp_hvc$term %in% bp_good,]
bp_mvc_filt <- bp_mvc[bp_mvc$term %in% bp_good,]
```

There are 5729 shared GO terms between the High v Control and Mid v Control comparisons

Combine sets 
```{r}
ress <- merge(bp_hvc_filt, bp_mvc_filt, by = "term")
```

The x is High v Control and the y is Mid v Control.

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

99 shared BP GO terms are < 0.01.

Fit a linear model to find the slope 
```{r}
# Fit a linear model
model <- lm(delta.rank.y ~ delta.rank.x, data = ress_filt)

# Extract the slope
slope <- coef(model)[2]

print(paste("The slope is:", round(slope, 4)))
```

Fit a rank regression to find the slope 
```{r}
## using csranks package
lmr_model <- lmranks(r(delta.rank.y) ~ r(delta.rank.x), data=ress_filt)
summary(lmr_model)

slope_rank <- coef(lmr_model)[2]
```

Create string for enriched delta ranks - these are the delta ranks that have the highest positive value in both comparisons 
```{r}
enrich_list <- c("regulation of interferon-alpha production", "DNA-templated DNA replication maintenance of fidelity", "DNA replication initiation", "recombinational repair", "DNA recombination")
enrich_list_label <- paste(enrich_list, collapse = "\n")
# Create a data frame for the label position
enrich_label_data <- data.frame(
  x = 600, # Adjust x position as needed
  y = 3100, # Adjust y position as needed
  label = enrich_list_label
)
```

Create string for underrepresented delta ranks - these are the delta ranks that have the most negative value in both comparisons 
```{r}
under_list <- c("mitochondrial ATP synthesis coupled electron transport", "oxidative phosphorylation", "electron transport chain", "cellular respiration", "generation of precursor metabolites and energy")
under_list_label <- paste(under_list, collapse = "\n")
# Create a data frame for the label position
under_label_data <- data.frame(
  x = -900, # Adjust x position as needed
  y = -2000, # Adjust y position as needed
  label = under_list_label
)
```

Color points to be labeled on plot 
```{r}
ress_filt <- ress_filt %>%
  mutate(include = case_when(
    str_detect(name.x, "regulation of interferon-alpha production|DNA-templated DNA replication maintenance of fidelity|DNA replication initiation|recombinational repair|DNA recombination") |
    str_detect(name.y, "regulation of interferon-alpha production|DNA-templated DNA replication maintenance of fidelity|DNA replication initiation|recombinational repair|DNA recombination") ~ "enriched",
    str_detect(name.x, "mitochondrial ATP synthesis coupled electron transport|oxidative phosphorylation|electron transport chain|cellular respiration|generation of precursor metabolites and energy") |
    str_detect(name.y, "mitochondrial ATP synthesis coupled electron transport|oxidative phosphorylation|electron transport chain|cellular respiration|generation of precursor metabolites and energy") ~ "under",
    
    TRUE ~ NA_character_
  ))
```

Plot 
```{r}
# Plot
plot3 <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y, color = include)) +
    # Negative horizontal dashed line (blue)
  geom_segment(aes(x = -Inf, xend = 0, y = 0, yend = 0), linetype = "dashed", alpha = 0.95, size = 1, color = "blue") +
  
  # Positive horizontal dashed line (red)
  geom_segment(aes(x = 0, xend = Inf, y = 0, yend = 0), linetype = "dashed", alpha = 0.95, size = 1, color = "red") +
  
  # Negative vertical dashed line (blue)
  geom_segment(aes(x = 0, xend = 0, y = -Inf, yend = 0), linetype = "dashed", alpha = 0.95, size = 1, color = "blue") +
  
  # Positive vertical dashed line (red)
  geom_segment(aes(x = 0, xend = 0, y = 0, yend = Inf), linetype = "dashed", alpha = 0.95, size = 1, color = "red") +
  
  geom_point(size = 4, show.legend = FALSE) +
  scale_size("size") +
  scale_x_continuous(limits = c(-3700, 3700), breaks = seq(-3000, 3000, 1000)) +
  scale_y_continuous(limits = c(-3700, 3700), breaks = seq(-3000, 3000, 1000)) +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (Mid v. Control)") +
  #geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.75) +
  #geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.75) +
  scale_color_manual(values = c("red", "blue", "black")) +
  # geom_label_repel(
  #   data = enrich_label_data,
  #   aes(x = x, y = y, label = label),
  #   box.padding = unit(1, "lines"),
  #   point.padding = unit(0.5, "lines"),
  #   size = 2.5,
  #   color = "red", # Text color
  #   fill = "white", # Background color of the label
  #   segment.color = NA) + # No line segment needed here
  # geom_label_repel(
  #   data = under_label_data,
  #   aes(x = x, y = y, label = label),
  #   box.padding = unit(1, "lines"),
  #   point.padding = unit(0.5, "lines"),
  #   size = 2.5,
  #   color = "blue", # Text color
  #   fill = "white", # Background color of the label
  #   segment.color = NA) + # No line segment needed here
  theme_cowplot() +
  annotate("label", label = "Overrepresented\nin High + Mid", x = 2600, y = 500, size = 7, color = "red", label.size = 0.8)+
  annotate("label", label = "Underrepresented\nin High + Mid", x = -2600, y = -500, size = 7, color = "blue", label.size = 0.8)+
  annotate("text", label = paste("Slope:", round(slope_rank, 2)), x = 2300, y = -3600, size = 6, color = "black") +
  theme(axis.text = element_text(size = 25),
        axis.title = element_text(size = 28)); plot3

ggsave("../output/GO_MWU/HighvControl_MidvControl_BP_DeltaRank.pdf", plot = plot3, width = 30, height = 25, units = "cm")
ggsave("../output/GO_MWU/HighvControl_MidvControl_BP_DeltaRank.png", plot = plot3, width = 30, height = 25, units = "cm")
```

#### Molecular Function 

Read in MF data. 
```{r}
mf_hvc=read.table("GO_MWU/MWU_MF_DE_High_v_Control_LFC.csv",header=T)
mf_mvc=read.table("GO_MWU/MWU_MF_DE_Mid_v_Control_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
mf_good <- intersect(mf_hvc$term, mf_mvc$term)

# Filter sets
mf_hvc_filt <- mf_hvc[mf_hvc$term %in% mf_good,]
mf_mvc_filt <- mf_mvc[mf_mvc$term %in% mf_good,]
```

There are 1107 shared GO terms between the High v Control and Mid v Control comparisons

Combine sets 
```{r}
ress <- merge(mf_hvc_filt, mf_mvc_filt, by = "term")
```

The x is High v Control and the y is Mid v Control  

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

23 shared MF GO terms are < 0.01.

Fit a linear model to find the slope 
```{r}
# Fit a linear model
model <- lm(delta.rank.y ~ delta.rank.x, data = ress_filt)

# Extract the slope
slope <- coef(model)[2]

print(paste("The slope is:", round(slope, 4)))
```

Fit a rank regression to find the slope 
```{r}
## using csranks package
lmr_model <- lmranks(r(delta.rank.y) ~ r(delta.rank.x), data=ress_filt)
summary(lmr_model)

slope_rank <- coef(lmr_model)[2]
```

Create string for enriched delta ranks - these are the delta ranks that have the highest positive value in both comparisons 
```{r}
enrich_list <- c("histone modifying activity", "histone binding", "DNA binding", "chromatin binding", "catalytic activity, acting on a nucleic acid")
enrich_list_label <- paste(enrich_list, collapse = "\n")
# Create a data frame for the label position
enrich_label_data <- data.frame(
  x = 500, # Adjust x position as needed
  y = 2000, # Adjust y position as needed
  label = enrich_list_label
)
```

Create string for underrepresented delta ranks - these are the delta ranks that have the most negative value in both comparisons 
```{r}
under_list <- c("oxidoreduction-driven active transmembrane transporter activity", "electron transfer activity", "proton transmembrane transporter activity", "primary active transmembrane transporter activity", "organic anion transmembrane transporter activity")
under_list_label <- paste(under_list, collapse = "\n")
# Create a data frame for the label position
under_label_data <- data.frame(
  x = -500, # Adjust x position as needed
  y = -2100, # Adjust y position as needed
  label = under_list_label
)
```

Color points to be labeled on plot 
```{r}
ress_filt <- ress_filt %>%
  mutate(include = case_when(
    str_detect(name.x, "histone modifying activity|histone binding|DNA binding|chromatin binding|	
catalytic activity, acting on a nucleic acid") |
    str_detect(name.y, "histone modifying activity|histone binding|DNA binding|chromatin binding|catalytic activity, acting on a nucleic acid") ~ "enriched",
    str_detect(name.x, "oxidoreduction-driven active transmembrane transporter activity|electron transfer activity|proton transmembrane transporter activity|primary active transmembrane transporter activity|organic anion transmembrane transporter activity") |
    str_detect(name.y, "oxidoreduction-driven active transmembrane transporter activity|electron transfer activity|proton transmembrane transporter activity|primary active transmembrane transporter activity|organic anion transmembrane transporter activity") ~ "under",
    
    TRUE ~ NA_character_
  ))
```


Plot 
```{r}
# Plot
plot4 <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y, color = include)) +
    # Negative horizontal dashed line (blue)
  geom_segment(aes(x = -Inf, xend = 0, y = 0, yend = 0), linetype = "dashed", alpha = 0.95, size = 1, color = "blue") +
  
  # Positive horizontal dashed line (red)
  geom_segment(aes(x = 0, xend = Inf, y = 0, yend = 0), linetype = "dashed", alpha = 0.95, size = 1, color = "red") +
  
  # Negative vertical dashed line (blue)
  geom_segment(aes(x = 0, xend = 0, y = -Inf, yend = 0), linetype = "dashed", alpha = 0.95, size = 1, color = "blue") +
  
  # Positive vertical dashed line (red)
  geom_segment(aes(x = 0, xend = 0, y = 0, yend = Inf), linetype = "dashed", alpha = 0.95, size = 1, color = "red") +
  
  geom_point(size = 4, show.legend = FALSE) +
  scale_size("size") +
  scale_x_continuous(limits = c(-3700, 3700), breaks = seq(-3000, 3000, 1000)) +
  scale_y_continuous(limits = c(-3700, 3700), breaks = seq(-3000, 3000, 1000)) +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (Mid v. Control)") +
  #geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.75) +
  #geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.75) +
  scale_color_manual(values = c("red", "blue", "black")) +
  # geom_label_repel(
  #   data = enrich_label_data,
  #   aes(x = x, y = y, label = label),
  #   box.padding = unit(1, "lines"),
  #   point.padding = unit(0.5, "lines"),
  #   size = 2.5,
  #   color = "red", # Text color
  #   fill = "white", # Background color of the label
  #   segment.color = NA) + # No line segment needed here
  # geom_label_repel(
  #   data = under_label_data,
  #   aes(x = x, y = y, label = label),
  #   box.padding = unit(1, "lines"),
  #   point.padding = unit(0.5, "lines"),
  #   size = 2.5,
  #   color = "blue", # Text color
  #   fill = "white", # Background color of the label
  #  segment.color = NA) + # No line segment needed here
  theme_cowplot()+
  annotate("label", label = "Overrepresented\nin High + Mid", x = 2600, y = 500, size = 7, color = "red", label.size = 0.8)+
  annotate("label", label = "Underrepresented\nin High + Mid", x = -2600, y = -500, size = 7, color = "blue", label.size = 0.8)+
  annotate("text", label = paste("Slope:", round(slope_rank, 2)), x = 2300, y = -3600, size = 6, color = "black") +
  theme(axis.text = element_text(size = 25),
        axis.title = element_text(size = 28)); plot4

ggsave("../output/GO_MWU/HighvControl_MidvControl_MF_DeltaRank.pdf", plot = plot4, width = 30, height = 25, units = "cm")
ggsave("../output/GO_MWU/HighvControl_MidvControl_MF_DeltaRank.png", plot = plot4, width = 30, height = 25, units = "cm")
```

Make plots into one figure and save 
```{r}
plot <- plot_grid(plot3, plot4, labels = c('A', 'B'))

ggsave("../output/GO_MWU/HighvMid_MidvControl_DeltaRank.pdf", plot = plot, width = 35, height = 15, units = "cm")
ggsave("../output/GO_MWU/HighvMid_MidvControl_DeltaRank.png", plot = plot, width = 35, height = 15, units = "cm")
```

Plot all the things 
```{r}
patchwork <- (plot1 + plot2) / (plot3 + plot4)
plot_all <- patchwork +
  plot_annotation(
    tag_levels = 'A',
    title = " Biological Processes                                                                                         Molecular Functions",
    theme = theme(
      plot.title = element_text(size = 25, hjust = 0.5)
    )
  ) & 
  theme(plot.tag = element_text(size = 30)); plot_all

ggsave("../output/GO_MWU/DeltaRank.pdf", plot = plot_all, width = 58, height = 50, units = "cm")
ggsave("../output/GO_MWU/DeltaRank.png", plot = plot_all, width = 58, height = 50, units = "cm", dpi = 700)
ggsave("../figures/FIGURE_5.PNG", plot = plot_all, width = 58, height = 50, units = "cm", dpi = 700)
```

```{r}
plot_all <- patchwork +
  plot_annotation(
    tag_levels = 'A',
    title = " Biological Processes                                                                                         Molecular Functions",
    theme = theme(
      plot.title = element_text(size = 25, hjust = 0.5)
    )
  ) & 
  theme(plot.tag = element_text(size = 60)); plot_all


patchwork + 
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 50))
```





Bind all dfs into one df and save for supplementary table
```{r}
# Add extra columns for ontology and comparisons 
bp_hvc$comparison <- "High_v_Control"
bp_hvc$ontology <- "Biological Processes"
mf_hvc$comparison <- "High_v_Control"
mf_hvc$ontology <- "Molecular Functions"

bp_hvm$comparison <- "High_v_Mid"
bp_hvm$ontology <- "Biological Processes"
mf_hvm$comparison <- "High_v_Mid"
mf_hvm$ontology <- "Molecular Functions"

bp_mvc$comparison <- "Mid_v_Control"
bp_mvc$ontology <- "Biological Processes"
mf_mvc$comparison <- "Mid_v_Control"
mf_mvc$ontology <- "Molecular Functions"

# Bind and save 
all_data <- rbind(bp_hvc, mf_hvc, bp_hvm, mf_hvm, bp_mvc, mf_mvc)
write.csv(all_data, "../output/GO_MWU/goMWU_results.csv")
```

