---
title: "GO MWU plotting"
author: "Jill Ashey"
date: "2024-07-23"
output: html_document
---

This script will plot the outputs from GO MWU. Specifically, I'm interested in plotting the delta rank of shared GO terms of each treatment comparison against one another. This will tell us if the GO terms that are shared across treatment comparisons have similar expression patterns. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggrepel)
library(cowplot)
library(kableExtra)
library(BiocManager)
#library("GOSim")
#library(rrvgo)
#library(org.Ce.eg.db)
```

### Delta rank of the shared GO terms between High v Control and High v Mid. 

#### Biological Processes 

Read in BP data. 
```{r}
bp_hvc=read.table("GO_MWU/MWU_BP_DE_High_v_Control_LFC.csv",header=T)
bp_hvm=read.table("GO_MWU/MWU_BP_DE_High_v_Mid_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
bp_good <- intersect(bp_hvc$term, bp_hvm$term)

# Filter sets
bp_hvc_filt <- bp_hvc[bp_hvc$term %in% bp_good,]
bp_hvm_filt <- bp_hvm[bp_hvm$term %in% bp_good,]
```

There are 5751 shared GO terms between the High v Control and High v Mid comparisons

Combine sets 
```{r}
ress <- merge(bp_hvc_filt, bp_hvm_filt, by = "term")
```

The x is High v Control and the y is High v Mid. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

106 shared BP GO terms are < 0.01.

Plot to assess delta rank correlation between shared BP GO terms
```{r}
# bp_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
#   geom_point(size = 2.5, show.legend = F) +
#   geom_text_repel(data = ress_filt, aes(), segment.alpha = 1, box.padding = 1, direction = "both", max.overlaps = Inf, size = 3, force = 3) +
#   scale_size("size") +
#   labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (High v. Mid)") +
#   geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
#   geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
#   theme_cowplot(); bp_plot

# Adjust text size and increase repulsion
bp_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = FALSE) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (High v. Mid)") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.75) +
  theme_cowplot(); bp_plot
```

#### Cellular Components 

Read in CC data. 
```{r}
cc_hvc=read.table("GO_MWU/MWU_CC_DE_High_v_Control_LFC.csv",header=T)
cc_hvm=read.table("GO_MWU/MWU_CC_DE_High_v_Mid_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
cc_good <- intersect(cc_hvc$term, cc_hvm$term)

# Filter sets
cc_hvc_filt <- cc_hvc[cc_hvc$term %in% cc_good,]
cc_hvm_filt <- cc_hvm[cc_hvm$term %in% cc_good,]
```

There are 804 shared GO terms between the High v Control and High v Mid comparisons

Combine sets 
```{r}
ress <- merge(cc_hvc_filt, cc_hvm_filt, by = "term")
```

The x is High v Control and the y is High v Mid. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

42 shared CC GO terms are < 0.01.

Plot to assess delta rank correlation between shared CC GO terms
```{r}
cc_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = F) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (High v. Mid)") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  theme_cowplot(); cc_plot
```

#### Molecular Function 

Read in MF data. 
```{r}
mf_hvc=read.table("GO_MWU/MWU_MF_DE_High_v_Control_LFC.csv",header=T)
mf_hvm=read.table("GO_MWU/MWU_MF_DE_High_v_Mid_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
mf_good <- intersect(mf_hvc$term, mf_hvm$term)

# Filter sets
mf_hvc_filt <- mf_hvc[mf_hvc$term %in% mf_good,]
mf_hvm_filt <- mf_hvm[mf_hvm$term %in% mf_good,]
```

There are 1105 shared GO terms between the High v Control and High v Mid comparisons

Combine sets 
```{r}
ress <- merge(mf_hvc_filt, mf_hvm_filt, by = "term")
```

The x is High v Control and the y is High v Mid. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

28 shared MF GO terms are < 0.01.

Plot to assess delta rank correlation between shared MF GO terms
```{r}
mf_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = F) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (High v. Mid)") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  theme_cowplot(); mf_plot
```

Make plots into one figure and save 
```{r}
plot <- plot_grid(bp_plot, mf_plot, labels = c('A', 'B'))

ggsave("../output/GO_MWU/HighvControl_HighvMid_DeltaRank.pdf", plot = plot, width = 50, height = 35, units = "cm")
ggsave("../output/GO_MWU/HighvControl_HighvMid_DeltaRank.png", plot = plot, width = 50, height = 35, units = "cm")
```


### Delta rank of the shared GO terms between High v Control and Mid v Control. 

#### Biological Processes 

Read in BP data. 
```{r}
bp_hvc=read.table("GO_MWU/MWU_BP_DE_High_v_Control_LFC.csv",header=T)
bp_mvc=read.table("GO_MWU/MWU_BP_DE_Mid_v_Control_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
bp_good <- intersect(bp_hvc$term, bp_mvc$term)

# Filter sets
bp_hvc_filt <- bp_hvc[bp_hvc$term %in% bp_good,]
bp_mvc_filt <- bp_mvc[bp_mvc$term %in% bp_good,]
```

There are 5729 shared GO terms between the High v Control and Mid v Control comparisons

Combine sets 
```{r}
ress <- merge(bp_hvc_filt, bp_mvc_filt, by = "term")
```

The x is High v Control and the y is Mid v Control.

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

99 shared BP GO terms are < 0.01.

Plot to assess delta rank correlation between shared BP GO terms
```{r}
bp_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = F) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (Mid v. Control)") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  theme_cowplot(); bp_plot
```

#### Cellular Components 

Read in CC data. 
```{r}
cc_hvc=read.table("GO_MWU/MWU_CC_DE_High_v_Control_LFC.csv",header=T)
cc_mvc=read.table("GO_MWU/MWU_CC_DE_Mid_v_Control_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
cc_good <- intersect(cc_hvc$term, cc_mvc$term)

# Filter sets
cc_hvc_filt <- cc_hvc[cc_hvc$term %in% cc_good,]
cc_mvc_filt <- cc_mvc[cc_mvc$term %in% cc_good,]
```

There are 802 shared GO terms between the High v Control and Mid v Control comparisons

Combine sets 
```{r}
ress <- merge(cc_hvc_filt, cc_mvc_filt, by = "term")
```

The x is High v Control and the y is Mid v Control. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

43 shared CC GO terms are < 0.01.

Plot to assess delta rank correlation between shared CC GO terms
```{r}
cc_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = F) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (Mid v. Control)") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  theme_cowplot(); cc_plot
```

#### Molecular Function 

Read in MF data. 
```{r}
mf_hvc=read.table("GO_MWU/MWU_MF_DE_High_v_Control_LFC.csv",header=T)
mf_mvc=read.table("GO_MWU/MWU_MF_DE_Mid_v_Control_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
mf_good <- intersect(mf_hvc$term, mf_mvc$term)

# Filter sets
mf_hvc_filt <- mf_hvc[mf_hvc$term %in% mf_good,]
mf_mvc_filt <- mf_mvc[mf_mvc$term %in% mf_good,]
```

There are 1107 shared GO terms between the High v Control and Mid v Control comparisons

Combine sets 
```{r}
ress <- merge(mf_hvc_filt, mf_mvc_filt, by = "term")
```

The x is High v Control and the y is Mid v Control  

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

23 shared MF GO terms are < 0.01.

Plot to assess delta rank correlation between shared MF GO terms
```{r}
mf_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = F) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (Mid v. Control)") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  theme_cowplot(); mf_plot
```

Make plots into one figure and save 
```{r}
plot <- plot_grid(bp_plot, mf_plot, labels = c('A', 'B'))

ggsave("../output/GO_MWU/HighvControl_MidvControl_DeltaRank.pdf", plot = plot, width = 50, height = 35, units = "cm")
ggsave("../output/GO_MWU/HighvControl_MidvControl_DeltaRank.png", plot = plot, width = 50, height = 35, units = "cm")
```



### Delta rank of the shared GO terms between High v Mid and Mid v Control. 

#### Biological Processes 

Read in BP data. 
```{r}
bp_hvm=read.table("GO_MWU/MWU_BP_DE_High_v_Mid_LFC.csv",header=T)
bp_mvc=read.table("GO_MWU/MWU_BP_DE_Mid_v_Control_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
bp_good <- intersect(bp_hvm$term, bp_mvc$term)

# Filter sets
bp_hvm_filt <- bp_hvm[bp_hvm$term %in% bp_good,]
bp_mvc_filt <- bp_mvc[bp_mvc$term %in% bp_good,]
```

There are 5853 shared GO terms between the High v Mid and Mid v Control comparisons

Combine sets 
```{r}
ress <- merge(bp_hvm_filt, bp_mvc_filt, by = "term")
```

The x is High v Mid and the y is Mid v Control. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

52 shared BP GO terms are < 0.01.

Plot to assess delta rank correlation between shared BP GO terms
```{r}
bp_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = F) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Mid)", y = "Delta Rank (Mid v. Control)") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  theme_cowplot(); bp_plot
```


#### Cellular Components 

Read in CC data. 
```{r}
cc_hvm=read.table("GO_MWU/MWU_CC_DE_High_v_Mid_LFC.csv",header=T)
cc_mvc=read.table("GO_MWU/MWU_CC_DE_Mid_v_Control_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
cc_good <- intersect(cc_hvm$term, cc_mvc$term)

# Filter sets
cc_hvm_filt <- cc_hvm[cc_hvm$term %in% cc_good,]
cc_mvc_filt <- cc_mvc[cc_mvc$term %in% cc_good,]
```

There are 809 shared GO terms between the High v Mid and Mid v Control comparisons

Combine sets 
```{r}
ress <- merge(cc_hvm_filt, cc_mvc_filt, by = "term")
```

The x is High v Mid and the y is Mid v Control. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

23 shared CC GO terms are < 0.01.

Plot to assess delta rank correlation between shared CC GO terms
```{r}
cc_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = F) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Mid)", y = "Delta Rank (Mid v. Control)") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  theme_cowplot(); cc_plot
```

#### Molecular Function 

Read in MF data. 
```{r}
mf_hvm=read.table("GO_MWU/MWU_MF_DE_High_v_Mid_LFC.csv",header=T)
mf_mvc=read.table("GO_MWU/MWU_MF_DE_Mid_v_Control_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
mf_good <- intersect(mf_hvm$term, mf_mvc$term)

# Filter sets
mf_hvm_filt <- mf_hvm[mf_hvm$term %in% mf_good,]
mf_mvc_filt <- mf_mvc[mf_mvc$term %in% mf_good,]
```

There are 1143 shared GO terms between the High v Mid and Mid v Control comparisons

Combine sets 
```{r}
ress <- merge(mf_hvm_filt, mf_mvc_filt, by = "term")
```

The x is High v Mid and the y is Mid v Control. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

19 shared MF GO terms are < 0.01.

Plot to assess delta rank correlation between shared CC GO terms
```{r}
mf_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = F) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Mid)", y = "Delta Rank (Mid v. Control)") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  theme_cowplot(); mf_plot
```

Make plots into one figure and save 
```{r}
plot <- plot_grid(bp_plot, mf_plot, labels = c('A', 'B'))

ggsave("../output/GO_MWU/HighvMid_MidvControl_DeltaRank.pdf", plot = plot, width = 50, height = 35, units = "cm")
ggsave("../output/GO_MWU/HighvMid_MidvControl_DeltaRank.png", plot = plot, width = 50, height = 35, units = "cm")
```

Make plot with comparisons for manuscript
```{r}
comparisons_table <- data.frame(
  Comparisons = c("High v. Control", "High v. Mid", "Mid v. Control"),
  "BiologicalProcesses" = c(583, 366, 551),
  "MolecularFunctions" = c(117, 96, 104)
)

# Set column names 
colnames(comparisons_table) <- c("Comparisons", "Biological Processes", "Molecular Functions")

# Create a formatted table using kable
kbl(comparisons_table) %>%
  kable_styling(full_width = F, position = "left") %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  column_spec(1, bold = T) %>%
  row_spec(0, bold = T) #%>%
  #save_kable("../output/GO_MWU/comparison_table.pdf")
```



Remaking manuscript figs so that there are not so many labels on the plots and each point isn't individually labeled 

## High v Control and High v Mid 

#### Biological Processes

Read in BP data. 
```{r}
bp_hvc=read.table("GO_MWU/MWU_BP_DE_High_v_Control_LFC.csv",header=T)
bp_hvm=read.table("GO_MWU/MWU_BP_DE_High_v_Mid_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
bp_good <- intersect(bp_hvc$term, bp_hvm$term)

# Filter sets
bp_hvc_filt <- bp_hvc[bp_hvc$term %in% bp_good,]
bp_hvm_filt <- bp_hvm[bp_hvm$term %in% bp_good,]
```

There are 5751 shared GO terms between the High v Control and High v Mid comparisons

Combine sets 
```{r}
ress <- merge(bp_hvc_filt, bp_hvm_filt, by = "term")
```

The x is High v Control and the y is High v Mid. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

106 shared BP GO terms are < 0.01.

Rank regression model - similar to linear regression but with no assumptions of normality or equal variance. It instead replaces the original data values with their corresponding ranks. This is more robust to outliers because it gives equal weight to all data regardless of magnitude
```{r}
# Create the rank regression model
x_rank <- rank(ress_filt$delta.rank.x)
y_rank <- rank(ress_filt$delta.rank.y)

rank_model <- lm(y_rank ~ x_rank)

# Look at the summary of the model
summary(rank_model)

# Create a scatter plot of the ranked data
plot(x_rank, y_rank, main="Rank Regression: Delta Rank Y vs Delta Rank X",
     xlab="Rank of Delta Rank X", ylab="Rank of Delta Rank Y")
abline(rank_model, col="red")

# OG data
plot(ress_filt$delta.rank.x, ress_filt$delta.rank.y, 
     main="Scatter Plot: Delta Rank Y vs Delta Rank X",
     xlab="Delta Rank X", ylab="Delta Rank Y")
```

Create string for enriched delta ranks - these are the delta ranks that have the highest positive value in both comparisons 
```{r}
enrich_list <- c("single strand break repair", "G-quadruplex DNA unwinding", "V(D)J recombination", "negative regulation of telomere capping", "mRNA 3’-splice recognition")
enrich_list_label <- paste(enrich_list, collapse = "\n")

# Create a data frame for the label position
enrich_label_data <- data.frame(
  x = 1000, # Adjust x position as needed
  y = 3100, # Adjust y position as needed
  label = enrich_list_label
)
```

Create string for underrepresented delta ranks - these are the delta ranks that have the most negative value in both comparisons 
```{r}
under_list <- c("lipoxygenase pathway", "axonemal dynein complex assembly", "long-chain fatty acid biosynthetic process", "cobalamin metabolic process", "fatty acid beta-oxidation", "NAD metabolic process")
under_list_label <- paste(under_list, collapse = "\n")

# Create a data frame for the label position
under_label_data <- data.frame(
  x = -4100, # Adjust x position as needed
  y = -600, # Adjust y position as needed
  label = under_list_label
)
```

Color points to be labeled on plot 
```{r}
ress_filt <- ress_filt %>%
  mutate(include = case_when(
    str_detect(name.x, "single strand break repair|G-quadruplex DNA unwinding|V\\(D\\)J recombination|negative regulation of telomere capping|mRNA.*3'-splice.*recognition") |
    str_detect(name.y, "single strand break repair|G-quadruplex DNA unwinding|V\\(D\\)J recombination|negative regulation of telomere capping|mRNA.*3'-splice.*recognition") ~ "enriched",
    str_detect(name.x, "lipoxygenase pathway|axonemal dynein complex assembly|long-chain fatty acid biosynthetic process|fatty acid beta-oxidation|NAD metabolic process|cobalamin metabolic process") |
    str_detect(name.y, "lipoxygenase pathway|axonemal dynein complex assembly|long-chain fatty acid biosynthetic process|fatty acid beta-oxidation|NAD metabolic process|cobalamin metabolic process") ~ "under",
    
    TRUE ~ NA_character_
  ))
```

Plot 
```{r}
# Plot
plot1 <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y, color = include)) +
  geom_point(size = 2, show.legend = FALSE) +
  scale_size("size") +
  #scale_x_continuous(limits = c(-2500, 2000), breaks = seq(-2000, 2000, 1000)) +
  #scale_y_continuous(limits = c(-2000, 2400), breaks = seq(-2000, 2000, 1000)) +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (High v. Mid)") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.75) +
  scale_color_manual(values = c("red", "blue", "black")) +
  geom_label_repel(
    data = enrich_label_data,
    aes(x = x, y = y, label = label),
    box.padding = unit(1, "lines"),
    point.padding = unit(0.5, "lines"),
    size = 2.5,
    color = "red", # Text color
    fill = "white", # Background color of the label
    segment.color = NA) + # No line segment needed here
  geom_label_repel(
    data = under_label_data,
    aes(x = x, y = y, label = label),
    box.padding = unit(1, "lines"),
    point.padding = unit(0.5, "lines"),
    size = 2.5,
    color = "blue", # Text color
    fill = "white", # Background color of the label
    segment.color = NA) + # No line segment needed here
  #annotate("text", label = "R^2 = 0.9702", x = 2200, y = -2300, size = 2, color = "black", parse = T)+
  #annotate("text", label = "p < 0.001", x = 2200, y = -2500, size = 2, color = "black", parse = T)+
  theme(axis.text = element_text(size = 50),
        axis.title = element_text(size = 70)) +
  theme_cowplot(); plot1

ggsave("../output/GO_MWU/HighvControl_HighvMid_BP_DeltaRank_boxlabel.pdf", plot = plot1, width = 20, height = 15, units = "cm")
ggsave("../output/GO_MWU/HighvControl_HighvMid_BP_DeltaRank_boxlabel.png", plot = plot1, width = 20, height = 15, units = "cm")
```



#### Molecular Function 

Read in MF data. 
```{r}
mf_hvc=read.table("GO_MWU/MWU_MF_DE_High_v_Control_LFC.csv",header=T)
mf_hvm=read.table("GO_MWU/MWU_MF_DE_High_v_Mid_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
mf_good <- intersect(mf_hvc$term, mf_hvm$term)

# Filter sets
mf_hvc_filt <- mf_hvc[mf_hvc$term %in% mf_good,]
mf_hvm_filt <- mf_hvm[mf_hvm$term %in% mf_good,]
```

There are 1105 shared GO terms between the High v Control and High v Mid comparisons

Combine sets 
```{r}
ress <- merge(mf_hvc_filt, mf_hvm_filt, by = "term")
```

The x is High v Control and the y is High v Mid. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

28 shared MF GO terms are < 0.01.

Rank regression model - similar to linear regression but with no assumptions of normality or equal variance. It instead replaces the original data values with their corresponding ranks. This is more robust to outliers because it gives equal weight to all data regardless of magnitude
```{r}
# Create the rank regression model
x_rank <- rank(ress_filt$delta.rank.x)
y_rank <- rank(ress_filt$delta.rank.y)

rank_model <- lm(y_rank ~ x_rank)

# Look at the summary of the model
summary(rank_model)

# Create a scatter plot of the ranked data
plot(x_rank, y_rank, main="Rank Regression: Delta Rank Y vs Delta Rank X",
     xlab="Rank of Delta Rank X", ylab="Rank of Delta Rank Y")
abline(rank_model, col="red")

# OG data
plot(ress_filt$delta.rank.x, ress_filt$delta.rank.y, 
     main="Scatter Plot: Delta Rank Y vs Delta Rank X",
     xlab="Delta Rank X", ylab="Delta Rank Y")
```

Create string for enriched delta ranks - these are the delta ranks that have the highest positive value in both comparisons 
```{r}
enrich_list <- c("single-stranded DNA helicase activity", "DNA phosphodiesterase activity", "histone kinase activity", "single-stranded DNA binding", "DNA helicase activity")
enrich_list_label <- paste(enrich_list, collapse = "\n")

# Create a data frame for the label position
enrich_label_data <- data.frame(
  x = 1000, # Adjust x position as needed
  y = 3300, # Adjust y position as needed
  label = enrich_list_label
)
```

Create string for underrepresented delta ranks - these are the delta ranks that have the most negative value in both comparisons 
```{r}
under_list <- c("oxidoreductase activity", "oxidoreductase activity, acting on single donors with incorporation of molecular oxygen")
under_list_label <- paste(under_list, collapse = "\n")

# Create a data frame for the label position
under_label_data <- data.frame(
  x = -3500, # Adjust x position as needed
  y = -1000, # Adjust y position as needed
  label = under_list_label
)
```

Color points to be labeled on plot 
```{r}
ress_filt <- ress_filt %>%
  mutate(include = case_when(
    str_detect(name.x, "single-stranded DNA helicase activity|DNA phosphodiesterase activity|histone kinase activity|single-stranded DNA binding|DNA helicase activity") |
    str_detect(name.y, "single-stranded DNA helicase activity|DNA phosphodiesterase activity|histone kinase activity|single-stranded DNA binding|DNA helicase activity") ~ "enriched",
    str_detect(name.x, "oxidoreductase activity|oxidoreductase activity, acting on single donors with incorporation of molecular oxygen") |
    str_detect(name.y, "oxidoreductase activity|oxidoreductase activity, acting on single donors with incorporation of molecular oxygen") ~ "under",
    
    TRUE ~ NA_character_
  ))
```

Plot 
```{r}
# Plot
plot2 <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y, color = include)) +
  geom_point(size = 2, show.legend = FALSE) +
  scale_size("size") +
  #scale_x_continuous(limits = c(-2500, 2000), breaks = seq(-2000, 2000, 1000)) +
  #scale_y_continuous(limits = c(-2000, 2400), breaks = seq(-2000, 2000, 1000)) +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (High v. Mid)") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.75) +
  scale_color_manual(values = c("red", "blue", "black")) +
  geom_label_repel(
    data = enrich_label_data,
    aes(x = x, y = y, label = label),
    box.padding = unit(1, "lines"),
    point.padding = unit(0.5, "lines"),
    size = 2.5,
    color = "red", # Text color
    fill = "white", # Background color of the label
    segment.color = NA) + # No line segment needed here
  geom_label_repel(
    data = under_label_data,
    aes(x = x, y = y, label = label),
    box.padding = unit(1, "lines"),
    point.padding = unit(0.5, "lines"),
    size = 2.5,
    color = "blue", # Text color
    fill = "white", # Background color of the label
    segment.color = NA) + # No line segment needed here
  #annotate("text", label = "R^2 = 0.8884", x = 2200, y = -2300, size = 2, color = "black", parse = T)+
  #annotate("text", label = "p < 0.001", x = 2200, y = -2500, size = 2, color = "black", parse = T)+
  theme(axis.text = element_text(size = 50),
        axis.title = element_text(size = 70)) +
  theme_cowplot(); plot2

ggsave("../output/GO_MWU/HighvControl_HighvMid_MF_DeltaRank_boxlabel.pdf", plot = plot2, width = 20, height = 15, units = "cm")
ggsave("../output/GO_MWU/HighvControl_HighvMid_MF_DeltaRank_boxlabel.png", plot = plot2, width = 20, height = 15, units = "cm")
```

Make plots into one figure and save 
```{r}
plot <- plot_grid(plot1, plot2, labels = c('A', 'B'))

ggsave("../output/GO_MWU/HighvMid_HighvControl_DeltaRank_boxlabel.pdf", plot = plot, width = 35, height = 15, units = "cm")
ggsave("../output/GO_MWU/HighvMid_HighvControl_DeltaRank_boxlabel.png", plot = plot, width = 35, height = 15, units = "cm")
```


## High v Control and Mid v Control

#### Biological Processes 

Read in BP data. 
```{r}
bp_hvc=read.table("GO_MWU/MWU_BP_DE_High_v_Control_LFC.csv",header=T)
bp_mvc=read.table("GO_MWU/MWU_BP_DE_Mid_v_Control_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
bp_good <- intersect(bp_hvc$term, bp_mvc$term)

# Filter sets
bp_hvc_filt <- bp_hvc[bp_hvc$term %in% bp_good,]
bp_mvc_filt <- bp_mvc[bp_mvc$term %in% bp_good,]
```

There are 5729 shared GO terms between the High v Control and Mid v Control comparisons

Combine sets 
```{r}
ress <- merge(bp_hvc_filt, bp_mvc_filt, by = "term")
```

The x is High v Control and the y is Mid v Control.

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

99 shared BP GO terms are < 0.01.

Rank regression model - similar to linear regression but with no assumptions of normality or equal variance. It instead replaces the original data values with their corresponding ranks. This is more robust to outliers because it gives equal weight to all data regardless of magnitude
```{r}
# Create the rank regression model
x_rank <- rank(ress_filt$delta.rank.x)
y_rank <- rank(ress_filt$delta.rank.y)

rank_model <- lm(y_rank ~ x_rank)

# Look at the summary of the model
summary(rank_model)

# Create a scatter plot of the ranked data
plot(x_rank, y_rank, main="Rank Regression: Delta Rank Y vs Delta Rank X",
     xlab="Rank of Delta Rank X", ylab="Rank of Delta Rank Y")
abline(rank_model, col="red")

# OG data
plot(ress_filt$delta.rank.x, ress_filt$delta.rank.y, 
     main="Scatter Plot: Delta Rank Y vs Delta Rank X",
     xlab="Delta Rank X", ylab="Delta Rank Y")
```

Create string for enriched delta ranks - these are the delta ranks that have the highest positive value in both comparisons 
```{r}
enrich_list <- c("regulation of interferon-alpha production", "DNA-templated DNA replication maintenance of fidelity", "DNA replication initiation", "recombinational repair", "DNA recombination")
enrich_list_label <- paste(enrich_list, collapse = "\n")

# Create a data frame for the label position
enrich_label_data <- data.frame(
  x = 500, # Adjust x position as needed
  y = 3100, # Adjust y position as needed
  label = enrich_list_label
)
```

Create string for underrepresented delta ranks - these are the delta ranks that have the most negative value in both comparisons 
```{r}
under_list <- c("mitochondrial ATP synthesis coupled electron transport", "oxidative phosphorylation", "electron transport chain", "cellular respiration", "generation of precursor metabolites and energy")
under_list_label <- paste(under_list, collapse = "\n")

# Create a data frame for the label position
under_label_data <- data.frame(
  x = -900, # Adjust x position as needed
  y = -2000, # Adjust y position as needed
  label = under_list_label
)
```

Color points to be labeled on plot 
```{r}
ress_filt <- ress_filt %>%
  mutate(include = case_when(
    str_detect(name.x, "regulation of interferon-alpha production|DNA-templated DNA replication maintenance of fidelity|DNA replication initiation|recombinational repair|DNA recombination") |
    str_detect(name.y, "regulation of interferon-alpha production|DNA-templated DNA replication maintenance of fidelity|DNA replication initiation|recombinational repair|DNA recombination") ~ "enriched",
    str_detect(name.x, "mitochondrial ATP synthesis coupled electron transport|oxidative phosphorylation|electron transport chain|cellular respiration|generation of precursor metabolites and energy") |
    str_detect(name.y, "mitochondrial ATP synthesis coupled electron transport|oxidative phosphorylation|electron transport chain|cellular respiration|generation of precursor metabolites and energy") ~ "under",
    
    TRUE ~ NA_character_
  ))
```

Plot 
```{r}
# Plot
plot3 <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y, color = include)) +
  geom_point(size = 2, show.legend = FALSE) +
  scale_size("size") +
  #scale_x_continuous(limits = c(-2500, 2000), breaks = seq(-2000, 2000, 1000)) +
  #scale_y_continuous(limits = c(-2000, 2400), breaks = seq(-2000, 2000, 1000)) +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (Mid v. Control)") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.75) +
  scale_color_manual(values = c("red", "blue", "black")) +
  geom_label_repel(
    data = enrich_label_data,
    aes(x = x, y = y, label = label),
    box.padding = unit(1, "lines"),
    point.padding = unit(0.5, "lines"),
    size = 2.5,
    color = "red", # Text color
    fill = "white", # Background color of the label
    segment.color = NA) + # No line segment needed here
  geom_label_repel(
    data = under_label_data,
    aes(x = x, y = y, label = label),
    box.padding = unit(1, "lines"),
    point.padding = unit(0.5, "lines"),
    size = 2.5,
    color = "blue", # Text color
    fill = "white", # Background color of the label
    segment.color = NA) + # No line segment needed here
  #annotate("text", label = "R^2 = 0.7201", x = 2200, y = -2300, size = 2, color = "black", parse = T)+
  #annotate("text", label = "p < 0.001", x = 2200, y = -2500, size = 2, color = "black", parse = T)+
  theme(axis.text = element_text(size = 50),
        axis.title = element_text(size = 70)) +
  theme_cowplot(); plot3

ggsave("../output/GO_MWU/HighvControl_MidvControl_BP_DeltaRank_boxlabel.pdf", plot = plot3, width = 20, height = 15, units = "cm")
ggsave("../output/GO_MWU/HighvControl_MidvControl_BP_DeltaRank_boxlabel.png", plot = plot3, width = 20, height = 15, units = "cm")
```

#### Molecular Function 

Read in MF data. 
```{r}
mf_hvc=read.table("GO_MWU/MWU_MF_DE_High_v_Control_LFC.csv",header=T)
mf_mvc=read.table("GO_MWU/MWU_MF_DE_Mid_v_Control_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
mf_good <- intersect(mf_hvc$term, mf_mvc$term)

# Filter sets
mf_hvc_filt <- mf_hvc[mf_hvc$term %in% mf_good,]
mf_mvc_filt <- mf_mvc[mf_mvc$term %in% mf_good,]
```

There are 1107 shared GO terms between the High v Control and Mid v Control comparisons

Combine sets 
```{r}
ress <- merge(mf_hvc_filt, mf_mvc_filt, by = "term")
```

The x is High v Control and the y is Mid v Control  

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

23 shared MF GO terms are < 0.01.

Rank regression model - similar to linear regression but with no assumptions of normality or equal variance. It instead replaces the original data values with their corresponding ranks. This is more robust to outliers because it gives equal weight to all data regardless of magnitude
```{r}
# Create the rank regression model
x_rank <- rank(ress_filt$delta.rank.x)
y_rank <- rank(ress_filt$delta.rank.y)

rank_model <- lm(y_rank ~ x_rank)

# Look at the summary of the model
summary(rank_model)

# Create a scatter plot of the ranked data
plot(x_rank, y_rank, main="Rank Regression: Delta Rank Y vs Delta Rank X",
     xlab="Rank of Delta Rank X", ylab="Rank of Delta Rank Y")
abline(rank_model, col="red")

# OG data
plot(ress_filt$delta.rank.x, ress_filt$delta.rank.y, 
     main="Scatter Plot: Delta Rank Y vs Delta Rank X",
     xlab="Delta Rank X", ylab="Delta Rank Y")
```

Even though the trend looks generally positive, the one point in the lower right quadrant is dragging down the R-squared value.

Create string for enriched delta ranks - these are the delta ranks that have the highest positive value in both comparisons 
```{r}
enrich_list <- c("histone modifying activity", "histone binding", "DNA binding", "chromatin binding", "catalytic activity, acting on a nucleic acid")
enrich_list_label <- paste(enrich_list, collapse = "\n")

# Create a data frame for the label position
enrich_label_data <- data.frame(
  x = 500, # Adjust x position as needed
  y = 2000, # Adjust y position as needed
  label = enrich_list_label
)
```

Create string for underrepresented delta ranks - these are the delta ranks that have the most negative value in both comparisons 
```{r}
under_list <- c("oxidoreduction-driven active transmembrane transporter activity", "electron transfer activity", "proton transmembrane transporter activity", "primary active transmembrane transporter activity", "organic anion transmembrane transporter activity")
under_list_label <- paste(under_list, collapse = "\n")

# Create a data frame for the label position
under_label_data <- data.frame(
  x = -500, # Adjust x position as needed
  y = -2100, # Adjust y position as needed
  label = under_list_label
)
```

Color points to be labeled on plot 
```{r}
ress_filt <- ress_filt %>%
  mutate(include = case_when(
    str_detect(name.x, "histone modifying activity|histone binding|DNA binding|chromatin binding|	
catalytic activity, acting on a nucleic acid") |
    str_detect(name.y, "histone modifying activity|histone binding|DNA binding|chromatin binding|catalytic activity, acting on a nucleic acid") ~ "enriched",
    str_detect(name.x, "oxidoreduction-driven active transmembrane transporter activity|electron transfer activity|proton transmembrane transporter activity|primary active transmembrane transporter activity|organic anion transmembrane transporter activity") |
    str_detect(name.y, "oxidoreduction-driven active transmembrane transporter activity|electron transfer activity|proton transmembrane transporter activity|primary active transmembrane transporter activity|organic anion transmembrane transporter activity") ~ "under",
    
    TRUE ~ NA_character_
  ))
```

Plot 
```{r}
# Plot
plot4 <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y, color = include)) +
  geom_point(size = 2, show.legend = FALSE) +
  scale_size("size") +
  #scale_x_continuous(limits = c(-2500, 2000), breaks = seq(-2000, 2000, 1000)) +
  #scale_y_continuous(limits = c(-2000, 2400), breaks = seq(-2000, 2000, 1000)) +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (Mid v. Control)") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.75) +
  scale_color_manual(values = c("red", "blue", "black")) +
  geom_label_repel(
    data = enrich_label_data,
    aes(x = x, y = y, label = label),
    box.padding = unit(1, "lines"),
    point.padding = unit(0.5, "lines"),
    size = 2.5,
    color = "red", # Text color
    fill = "white", # Background color of the label
    segment.color = NA) + # No line segment needed here
  geom_label_repel(
    data = under_label_data,
    aes(x = x, y = y, label = label),
    box.padding = unit(1, "lines"),
    point.padding = unit(0.5, "lines"),
    size = 2.5,
    color = "blue", # Text color
    fill = "white", # Background color of the label
    segment.color = NA) + # No line segment needed here
  #annotate("text", label = "R^2 = 0.197", x = 2200, y = -2300, size = 2, color = "black", parse = T)+
  #annotate("text", label = "p < 0.01", x = 2200, y = -2500, size = 2, color = "black", parse = T)+
  theme(axis.text = element_text(size = 50),
        axis.title = element_text(size = 70)) +
  theme_cowplot(); plot4

ggsave("../output/GO_MWU/HighvControl_MidvControl_MF_DeltaRank_boxlabel.pdf", plot = plot4, width = 20, height = 15, units = "cm")
ggsave("../output/GO_MWU/HighvControl_MidvControl_MF_DeltaRank_boxlabel.png", plot = plot4, width = 20, height = 15, units = "cm")
```

Make plots into one figure and save 
```{r}
plot <- plot_grid(plot3, plot4, labels = c('A', 'B'))

ggsave("../output/GO_MWU/HighvMid_MidvControl_DeltaRank_boxlabel.pdf", plot = plot, width = 35, height = 15, units = "cm")
ggsave("../output/GO_MWU/HighvMid_MidvControl_DeltaRank_boxlabel.png", plot = plot, width = 35, height = 15, units = "cm")
```

Plot all the things 
```{r}
plot_all <- plot_grid(plot1, plot2, plot3, plot4, labels = c('A', 'B', 'C', 'D'), nrow = 2, ncol = 2)

ggsave("../output/GO_MWU/DeltaRank_boxlabel.pdf", plot = plot_all, width = 40, height = 25, units = "cm")
ggsave("../output/GO_MWU/DeltaRank_boxlabel.png", plot = plot_all, width = 40, height = 25, units = "cm")
```

Bind all dfs into one df and save for supplementary table
```{r}
# Add extra columns for ontology and comparisons 
bp_hvc$comparison <- "High_v_Control"
bp_hvc$ontology <- "Biological Processes"
mf_hvc$comparison <- "High_v_Control"
mf_hvc$ontology <- "Molecular Functions"

bp_hvm$comparison <- "High_v_Mid"
bp_hvm$ontology <- "Biological Processes"
mf_hvm$comparison <- "High_v_Mid"
mf_hvm$ontology <- "Molecular Functions"

bp_mvc$comparison <- "Mid_v_Control"
bp_mvc$ontology <- "Biological Processes"
mf_mvc$comparison <- "Mid_v_Control"
mf_mvc$ontology <- "Molecular Functions"

# Bind and save 
all_data <- rbind(bp_hvc, mf_hvc, bp_hvm, mf_hvm, bp_mvc, mf_mvc)
write.csv(all_data, "../output/GO_MWU/goMWU_results.csv")
```

