---
title: "GO MWU plotting"
author: "Jill Ashey"
date: "2024-07-23"
output: html_document
---

This script will plot the outputs from GO MWU. Specifically, I'm interested in plotting the delta rank of shared GO terms of each treatment comparison against one another. This will tell us if the GO terms that are shared across treatment comparisons have similar expression patterns. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggrepel)
library(cowplot)
library(kableExtra)
library(BiocManager)
#library("GOSim")
library(rrvgo)
library(org.Ce.eg.db)
```

### Delta rank of the shared GO terms between High v Control and High v Mid. 

#### Biological Processes 

Read in BP data. 
```{r}
bp_hvc=read.table("GO_MWU/MWU_BP_DE_High_v_Control_LFC.csv",header=T)
bp_hvm=read.table("GO_MWU/MWU_BP_DE_High_v_Mid_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
bp_good <- intersect(bp_hvc$term, bp_hvm$term)

# Filter sets
bp_hvc_filt <- bp_hvc[bp_hvc$term %in% bp_good,]
bp_hvm_filt <- bp_hvm[bp_hvm$term %in% bp_good,]
```

There are 5751 shared GO terms between the High v Control and High v Mid comparisons

Combine sets 
```{r}
ress <- merge(bp_hvc_filt, bp_hvm_filt, by = "term")
```

The x is High v Control and the y is High v Mid. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

106 shared BP GO terms are < 0.01.

Plot to assess delta rank correlation between shared BP GO terms
```{r}
# bp_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
#   geom_point(size = 2.5, show.legend = F) +
#   geom_text_repel(data = ress_filt, aes(), segment.alpha = 1, box.padding = 1, direction = "both", max.overlaps = Inf, size = 3, force = 3) +
#   scale_size("size") +
#   labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (High v. Mid)") +
#   geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
#   geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
#   theme_cowplot(); bp_plot

# Adjust text size and increase repulsion
bp_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = FALSE) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (High v. Mid)") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.75) +
  theme_cowplot(); bp_plot
```

#### Cellular Components 

Read in CC data. 
```{r}
cc_hvc=read.table("GO_MWU/MWU_CC_DE_High_v_Control_LFC.csv",header=T)
cc_hvm=read.table("GO_MWU/MWU_CC_DE_High_v_Mid_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
cc_good <- intersect(cc_hvc$term, cc_hvm$term)

# Filter sets
cc_hvc_filt <- cc_hvc[cc_hvc$term %in% cc_good,]
cc_hvm_filt <- cc_hvm[cc_hvm$term %in% cc_good,]
```

There are 804 shared GO terms between the High v Control and High v Mid comparisons

Combine sets 
```{r}
ress <- merge(cc_hvc_filt, cc_hvm_filt, by = "term")
```

The x is High v Control and the y is High v Mid. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

42 shared CC GO terms are < 0.01.

Plot to assess delta rank correlation between shared CC GO terms
```{r}
cc_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = F) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (High v. Mid)") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  theme_cowplot(); cc_plot
```

#### Molecular Function 

Read in MF data. 
```{r}
mf_hvc=read.table("GO_MWU/MWU_MF_DE_High_v_Control_LFC.csv",header=T)
mf_hvm=read.table("GO_MWU/MWU_MF_DE_High_v_Mid_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
mf_good <- intersect(mf_hvc$term, mf_hvm$term)

# Filter sets
mf_hvc_filt <- mf_hvc[mf_hvc$term %in% mf_good,]
mf_hvm_filt <- mf_hvm[mf_hvm$term %in% mf_good,]
```

There are 1105 shared GO terms between the High v Control and High v Mid comparisons

Combine sets 
```{r}
ress <- merge(mf_hvc_filt, mf_hvm_filt, by = "term")
```

The x is High v Control and the y is High v Mid. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

28 shared MF GO terms are < 0.01.

Plot to assess delta rank correlation between shared MF GO terms
```{r}
mf_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = F) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (High v. Mid)") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  theme_cowplot(); mf_plot
```

Make plots into one figure and save 
```{r}
plot <- plot_grid(bp_plot, mf_plot, labels = c('A', 'B'))

ggsave("../output/GO_MWU/HighvControl_HighvMid_DeltaRank.pdf", plot = plot, width = 50, height = 35, units = "cm")
ggsave("../output/GO_MWU/HighvControl_HighvMid_DeltaRank.png", plot = plot, width = 50, height = 35, units = "cm")
```


### Delta rank of the shared GO terms between High v Control and Mid v Control. 

#### Biological Processes 

Read in BP data. 
```{r}
bp_hvc=read.table("GO_MWU/MWU_BP_DE_High_v_Control_LFC.csv",header=T)
bp_mvc=read.table("GO_MWU/MWU_BP_DE_Mid_v_Control_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
bp_good <- intersect(bp_hvc$term, bp_mvc$term)

# Filter sets
bp_hvc_filt <- bp_hvc[bp_hvc$term %in% bp_good,]
bp_mvc_filt <- bp_mvc[bp_mvc$term %in% bp_good,]
```

There are 5729 shared GO terms between the High v Control and Mid v Control comparisons

Combine sets 
```{r}
ress <- merge(bp_hvc_filt, bp_mvc_filt, by = "term")
```

The x is High v Control and the y is Mid v Control.

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

99 shared BP GO terms are < 0.01.

Plot to assess delta rank correlation between shared BP GO terms
```{r}
bp_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = F) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (Mid v. Control)") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  theme_cowplot(); bp_plot
```

#### Cellular Components 

Read in CC data. 
```{r}
cc_hvc=read.table("GO_MWU/MWU_CC_DE_High_v_Control_LFC.csv",header=T)
cc_mvc=read.table("GO_MWU/MWU_CC_DE_Mid_v_Control_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
cc_good <- intersect(cc_hvc$term, cc_mvc$term)

# Filter sets
cc_hvc_filt <- cc_hvc[cc_hvc$term %in% cc_good,]
cc_mvc_filt <- cc_mvc[cc_mvc$term %in% cc_good,]
```

There are 802 shared GO terms between the High v Control and Mid v Control comparisons

Combine sets 
```{r}
ress <- merge(cc_hvc_filt, cc_mvc_filt, by = "term")
```

The x is High v Control and the y is Mid v Control. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

43 shared CC GO terms are < 0.01.

Plot to assess delta rank correlation between shared CC GO terms
```{r}
cc_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = F) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (Mid v. Control)") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  theme_cowplot(); cc_plot
```

#### Molecular Function 

Read in MF data. 
```{r}
mf_hvc=read.table("GO_MWU/MWU_MF_DE_High_v_Control_LFC.csv",header=T)
mf_mvc=read.table("GO_MWU/MWU_MF_DE_Mid_v_Control_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
mf_good <- intersect(mf_hvc$term, mf_mvc$term)

# Filter sets
mf_hvc_filt <- mf_hvc[mf_hvc$term %in% mf_good,]
mf_mvc_filt <- mf_mvc[mf_mvc$term %in% mf_good,]
```

There are 1107 shared GO terms between the High v Control and Mid v Control comparisons

Combine sets 
```{r}
ress <- merge(mf_hvc_filt, mf_mvc_filt, by = "term")
```

The x is High v Control and the y is Mid v Control  

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

23 shared MF GO terms are < 0.01.

Plot to assess delta rank correlation between shared MF GO terms
```{r}
mf_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = F) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (Mid v. Control)") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  theme_cowplot(); mf_plot
```

Make plots into one figure and save 
```{r}
plot <- plot_grid(bp_plot, mf_plot, labels = c('A', 'B'))

ggsave("../output/GO_MWU/HighvControl_MidvControl_DeltaRank.pdf", plot = plot, width = 50, height = 35, units = "cm")
ggsave("../output/GO_MWU/HighvControl_MidvControl_DeltaRank.png", plot = plot, width = 50, height = 35, units = "cm")
```



### Delta rank of the shared GO terms between High v Mid and Mid v Control. 

#### Biological Processes 

Read in BP data. 
```{r}
bp_hvm=read.table("GO_MWU/MWU_BP_DE_High_v_Mid_LFC.csv",header=T)
bp_mvc=read.table("GO_MWU/MWU_BP_DE_Mid_v_Control_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
bp_good <- intersect(bp_hvm$term, bp_mvc$term)

# Filter sets
bp_hvm_filt <- bp_hvm[bp_hvm$term %in% bp_good,]
bp_mvc_filt <- bp_mvc[bp_mvc$term %in% bp_good,]
```

There are 5853 shared GO terms between the High v Mid and Mid v Control comparisons

Combine sets 
```{r}
ress <- merge(bp_hvm_filt, bp_mvc_filt, by = "term")
```

The x is High v Mid and the y is Mid v Control. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

52 shared BP GO terms are < 0.01.

Plot to assess delta rank correlation between shared BP GO terms
```{r}
bp_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = F) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Mid)", y = "Delta Rank (Mid v. Control)") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  theme_cowplot(); bp_plot
```


#### Cellular Components 

Read in CC data. 
```{r}
cc_hvm=read.table("GO_MWU/MWU_CC_DE_High_v_Mid_LFC.csv",header=T)
cc_mvc=read.table("GO_MWU/MWU_CC_DE_Mid_v_Control_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
cc_good <- intersect(cc_hvm$term, cc_mvc$term)

# Filter sets
cc_hvm_filt <- cc_hvm[cc_hvm$term %in% cc_good,]
cc_mvc_filt <- cc_mvc[cc_mvc$term %in% cc_good,]
```

There are 809 shared GO terms between the High v Mid and Mid v Control comparisons

Combine sets 
```{r}
ress <- merge(cc_hvm_filt, cc_mvc_filt, by = "term")
```

The x is High v Mid and the y is Mid v Control. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

23 shared CC GO terms are < 0.01.

Plot to assess delta rank correlation between shared CC GO terms
```{r}
cc_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = F) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Mid)", y = "Delta Rank (Mid v. Control)") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  theme_cowplot(); cc_plot
```

#### Molecular Function 

Read in MF data. 
```{r}
mf_hvm=read.table("GO_MWU/MWU_MF_DE_High_v_Mid_LFC.csv",header=T)
mf_mvc=read.table("GO_MWU/MWU_MF_DE_Mid_v_Control_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
mf_good <- intersect(mf_hvm$term, mf_mvc$term)

# Filter sets
mf_hvm_filt <- mf_hvm[mf_hvm$term %in% mf_good,]
mf_mvc_filt <- mf_mvc[mf_mvc$term %in% mf_good,]
```

There are 1143 shared GO terms between the High v Mid and Mid v Control comparisons

Combine sets 
```{r}
ress <- merge(mf_hvm_filt, mf_mvc_filt, by = "term")
```

The x is High v Mid and the y is Mid v Control. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

19 shared MF GO terms are < 0.01.

Plot to assess delta rank correlation between shared CC GO terms
```{r}
mf_plot <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y)) +
  geom_point(size = 2.5, show.legend = F) +
  geom_text_repel(
    segment.alpha = 1,
    box.padding = 0.7,
    direction = "both",
    max.overlaps = Inf, # Allow more overlaps but control with point.size
    size = 3, # Reduce text size
    force = 2, # Increase repulsion force
    max.iter = 1000 # Increase iterations for better placement
  ) +
  scale_size("size") +
  labs(x = "Delta Rank (High v. Mid)", y = "Delta Rank (Mid v. Control)") +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.75) +
  theme_cowplot(); mf_plot
```

Make plots into one figure and save 
```{r}
plot <- plot_grid(bp_plot, mf_plot, labels = c('A', 'B'))

ggsave("../output/GO_MWU/HighvMid_MidvControl_DeltaRank.pdf", plot = plot, width = 50, height = 35, units = "cm")
ggsave("../output/GO_MWU/HighvMid_MidvControl_DeltaRank.png", plot = plot, width = 50, height = 35, units = "cm")
```

Make plot with comparisons for manuscript
```{r}
comparisons_table <- data.frame(
  Comparisons = c("High v. Control", "High v. Mid", "Mid v. Control"),
  "BiologicalProcesses" = c(583, 366, 551),
  "MolecularFunctions" = c(117, 96, 104)
)

# Set column names 
colnames(comparisons_table) <- c("Comparisons", "Biological Processes", "Molecular Functions")

# Create a formatted table using kable
kbl(comparisons_table) %>%
  kable_styling(full_width = F, position = "left") %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  column_spec(1, bold = T) %>%
  row_spec(0, bold = T) #%>%
  #save_kable("../output/GO_MWU/comparison_table.pdf")
```



Trying to create a new plot 

Hollie ideas: put all 4 of the plots of the delta ranks together in a matrix form, and then either color code the dots if few functions, or we will have to make some lists of only the conserved + and - DR High vs Mid compared to High Vs Control 

```{r}
#Reduce/collapse GO term set with the rrvgo package 
      simMatrix <- calculateSimMatrix(bp_hvc_filt$term,
                                orgdb="org.Ce.eg.db", #c. elegans database
                                ont=c("BP"),
                                method="Rel")
```
^^ come back to this 


Remake plots for manuscript so that the x and y axes have the same number and the dots aren't labeled 


## High v Control and High v Mid 

#### Biological Processes

Read in BP data. 
```{r}
bp_hvc=read.table("GO_MWU/MWU_BP_DE_High_v_Control_LFC.csv",header=T)
bp_hvm=read.table("GO_MWU/MWU_BP_DE_High_v_Mid_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
bp_good <- intersect(bp_hvc$term, bp_hvm$term)

# Filter sets
bp_hvc_filt <- bp_hvc[bp_hvc$term %in% bp_good,]
bp_hvm_filt <- bp_hvm[bp_hvm$term %in% bp_good,]
```

There are 5751 shared GO terms between the High v Control and High v Mid comparisons

Combine sets 
```{r}
ress <- merge(bp_hvc_filt, bp_hvm_filt, by = "term")
```

The x is High v Control and the y is High v Mid. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

106 shared BP GO terms are < 0.01.

Create string for enriched delta ranks - these are the delta ranks that have the highest positive value in both comparisons 
```{r}
enrich_list <- c("single strand break repair", "G-quadruplex DNA unwinding", "chromosome separation", "RNA and mRNA 3’-splice recognition")
enrich_list_label <- paste(enrich_list, collapse = "\n")

# Create a data frame for the label position
enrich_label_data <- data.frame(
  x = 1000, # Adjust x position as needed
  y = 3100, # Adjust y position as needed
  label = enrich_list_label
)
```

Create string for underrepresented delta ranks - these are the delta ranks that have the most negative value in both comparisons 
```{r}
under_list <- c("lipoxygenase pathway", "axonemal dynein complex assembly", "long-chain fatty acid biosynthetic process", "fatty acid beta-oxidation", "NAD metabolic process")
under_list_label <- paste(under_list, collapse = "\n")

# Create a data frame for the label position
under_label_data <- data.frame(
  x = -4100, # Adjust x position as needed
  y = -800, # Adjust y position as needed
  label = under_list_label
)
```

Color points to be labeled on plot 
```{r}
ress_filt <- ress_filt %>%
  mutate(include = case_when(
    str_detect(name.x, "single strand break repair|G-quadruplex DNA unwinding|chromosome separation|RNA.*3'-splice.*recognition|mRNA.*3'-splice.*recognition") |
    str_detect(name.y, "single strand break repair|G-quadruplex DNA unwinding|chromosome separation|RNA.*3'-splice.*recognition|mRNA.*3'-splice.*recognition") ~ "enriched",
    
    str_detect(name.x, "lipoxygenase pathway|axonemal dynein complex assembly|long-chain fatty acid biosynthetic process|fatty acid beta-oxidation|NAD metabolic process") |
    str_detect(name.y, "lipoxygenase pathway|axonemal dynein complex assembly|long-chain fatty acid biosynthetic process|fatty acid beta-oxidation|NAD metabolic process") ~ "over",
    
    TRUE ~ NA_character_
  ))
```

Plot 
```{r}
# Plot
plot1 <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y, color = include)) +
  geom_point(size = 2, show.legend = FALSE) +
  scale_size("size") +
  #scale_x_continuous(limits = c(-2500, 2000), breaks = seq(-2000, 2000, 1000)) +
  #scale_y_continuous(limits = c(-2000, 2400), breaks = seq(-2000, 2000, 1000)) +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (High v. Mid)") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.75) +
  scale_color_manual(values = c("red", "blue", "black")) +
  geom_label_repel(
    data = enrich_label_data,
    aes(x = x, y = y, label = label),
    box.padding = unit(1, "lines"),
    point.padding = unit(0.5, "lines"),
    size = 2.5,
    color = "black", # Text color
    fill = "white", # Background color of the label
    segment.color = NA) + # No line segment needed here
  geom_label_repel(
    data = under_label_data,
    aes(x = x, y = y, label = label),
    box.padding = unit(1, "lines"),
    point.padding = unit(0.5, "lines"),
    size = 2.5,
    color = "black", # Text color
    fill = "white", # Background color of the label
    segment.color = NA) + # No line segment needed here
  theme(axis.text = element_text(size = 50),
        axis.title = element_text(size = 70)) +
  theme_cowplot(); plot1

ggsave("../output/GO_MWU/HighvControl_HighvMid_BP_DeltaRank_boxlabel.pdf", plot = plot1, width = 20, height = 15, units = "cm")
ggsave("../output/GO_MWU/HighvControl_HighvMid_BP_DeltaRank_boxlabel.png", plot = plot1, width = 20, height = 15, units = "cm")
```



#### Molecular Function 

Read in MF data. 
```{r}
mf_hvc=read.table("GO_MWU/MWU_MF_DE_High_v_Control_LFC.csv",header=T)
mf_hvm=read.table("GO_MWU/MWU_MF_DE_High_v_Mid_LFC.csv",header=T)
```

Find terms in both sets and filter both sets by shared terms. 
```{r}
mf_good <- intersect(mf_hvc$term, mf_hvm$term)

# Filter sets
mf_hvc_filt <- mf_hvc[mf_hvc$term %in% mf_good,]
mf_hvm_filt <- mf_hvm[mf_hvm$term %in% mf_good,]
```

There are 1105 shared GO terms between the High v Control and High v Mid comparisons

Combine sets 
```{r}
ress <- merge(mf_hvc_filt, mf_hvm_filt, by = "term")
```

The x is High v Control and the y is High v Mid. 

Filter so that only p.adj < 0.01 for both sets
```{r}
ress_filt <- ress %>%
  dplyr::filter(p.adj.x < 0.01) %>%
  dplyr::filter(p.adj.y < 0.01) %>%
  dplyr::filter(!grepl("obsolete", name.x))
```

28 shared MF GO terms are < 0.01.

Create string for enriched delta ranks - these are the delta ranks that have the highest positive value in both comparisons 
```{r}
enrich_list <- c("single-stranded DNA helicase activity", "DNA phosphodiesterase activity", "histone kinase activity", "damaged DNA binding")
enrich_list_label <- paste(enrich_list, collapse = "\n")

# Create a data frame for the label position
enrich_label_data <- data.frame(
  x = 1000, # Adjust x position as needed
  y = 3100, # Adjust y position as needed
  label = enrich_list_label
)
```

Create string for underrepresented delta ranks - these are the delta ranks that have the most negative value in both comparisons 
```{r}
under_list <- c("oxidoreductase activity", "oxidoreductase activity, acting on single donors with incorporation of molecular oxygen")
under_list_label <- paste(under_list, collapse = "\n")

# Create a data frame for the label position
under_label_data <- data.frame(
  x = -3500, # Adjust x position as needed
  y = -1000, # Adjust y position as needed
  label = under_list_label
)
```

Color points to be labeled on plot 
```{r}
ress_filt <- ress_filt %>%
  mutate(include = case_when(
    str_detect(name.x, "single-stranded DNA helicase activity|DNA phosphodiesterase activity|histone kinase activity|damaged DNA binding") |
    str_detect(name.y, "single-stranded DNA helicase activity|DNA phosphodiesterase activity|histone kinase activity|damaged DNA binding") ~ "enriched",
    
    str_detect(name.x, "oxidoreductase activity|oxidoreductase activity, acting on single donors with incorporation of molecular oxygen") |
    str_detect(name.y, "oxidoreductase activity|oxidoreductase activity, acting on single donors with incorporation of molecular oxygen") ~ "over",
    
    TRUE ~ NA_character_
  ))
```

Plot 
```{r}
# Plot
plot2 <- ggplot(ress_filt, aes(delta.rank.x, delta.rank.y, label = name.y, color = include)) +
  geom_point(size = 2, show.legend = FALSE) +
  scale_size("size") +
  #scale_x_continuous(limits = c(-2500, 2000), breaks = seq(-2000, 2000, 1000)) +
  #scale_y_continuous(limits = c(-2000, 2400), breaks = seq(-2000, 2000, 1000)) +
  labs(x = "Delta Rank (High v. Control)", y = "Delta Rank (High v. Mid)") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.75) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.75) +
  scale_color_manual(values = c("red", "blue", "black")) +
  geom_label_repel(
    data = enrich_label_data,
    aes(x = x, y = y, label = label),
    box.padding = unit(1, "lines"),
    point.padding = unit(0.5, "lines"),
    size = 2.5,
    color = "black", # Text color
    fill = "white", # Background color of the label
    segment.color = NA) + # No line segment needed here
  geom_label_repel(
    data = under_label_data,
    aes(x = x, y = y, label = label),
    box.padding = unit(1, "lines"),
    point.padding = unit(0.5, "lines"),
    size = 2.5,
    color = "black", # Text color
    fill = "white", # Background color of the label
    segment.color = NA) + # No line segment needed here
  theme(axis.text = element_text(size = 50),
        axis.title = element_text(size = 70)) +
  theme_cowplot(); plot2

ggsave("../output/GO_MWU/HighvControl_HighvMid_MF_DeltaRank_boxlabel.pdf", plot = plot2, width = 20, height = 15, units = "cm")
ggsave("../output/GO_MWU/HighvControl_HighvMid_MF_DeltaRank_boxlabel.png", plot = plot2, width = 20, height = 15, units = "cm")
```

Make plots into one figure and save 
```{r}
plot <- plot_grid(plot1, plot2, labels = c('A', 'B'))

ggsave("../output/GO_MWU/HighvMid_HighvControl_DeltaRank_boxlabel.pdf", plot = plot, width = 35, height = 15, units = "cm")
ggsave("../output/GO_MWU/HighvMid_HighvControl_DeltaRank_boxlabel.png", plot = plot, width = 35, height = 15, units = "cm")
```


TO DO: make sure list of terms match those in manuscript, make other plots 


Rank regression model - similar to linear regression but with no assumptions of normality or equal variance. It instead replaces the original data values with their corresponding ranks. This is more robust to outliers because it gives equal weight to all data regardless of magnitude
```{r}
# Create the rank regression model
x_rank <- rank(ress_filt$delta.rank.x)
y_rank <- rank(ress_filt$delta.rank.y)

rank_model <- lm(y_rank ~ x_rank)

# Look at the summary of the model
summary(rank_model)

# Create a scatter plot of the ranked data
plot(x_rank, y_rank, main="Rank Regression: Delta Rank Y vs Delta Rank X",
     xlab="Rank of Delta Rank X", ylab="Rank of Delta Rank Y")
abline(rank_model, col="red")

# To visualize the original data
plot(ress_filt$delta.rank.x, ress_filt$delta.rank.y, 
     main="Scatter Plot: Delta Rank Y vs Delta Rank X",
     xlab="Delta Rank X", ylab="Delta Rank Y")
```

Do we need to include this information in the paper?






